\renewcommand{\abstractname}{Preface}
\begin{abstract}
\noindent
This thesis has been submitted in the fulfilment of Masters of Science in Communication Technology, with a specialization in information security at the \gls{ntnu} in Trondheim. The thesis is original, unpublished, and independent work by the author E. Klevstad.

%\noindent
%What you are about to read is the end of my road at \gls{ntnu}. A road which Google Maps estimated the time of arrival to be roughly five years. It has had 33 speed bumps (read exams), enormous fuel expenses (student loans), and a one-year long detour at the \gls{ucsb}, but the end of the road is finally here.


\noindent
I would like to thank my supervisor, Britta Hale, for her endeavours in the tweaking process of Scyther models, as well as valuable feedback and insight in the modelling of security protocols. I would also thank my responsible professor, Colin Boyd, for his valuable guidance, feedback, and support during this thesis. You have both been an incredible resource.

\noindent
Another round of ``thank yous'' goes to Google for developing Google Translate, and Britta and Colin for correcting Google Translate when it was wrong. A special shout-out to the people at my office for providing me with a fair amount of procrastination activities. I will treasure our time spent watching Norwegian slow TV and browsing the dark side of YouTube. Writing this thesis has been challenging from time to time. Therefore, I am forever grateful for the inspiration provided by you, Doppio Passo. I could not have done this without you.

\noindent
Finally, I would like to thank you as a reader. By completing these last sentences, you have at least read one page of my thesis.\\

\noindent
Cheers.\\

\noindent
\textbf{Eirik Klevstad}

\noindent
Trondheim,  \today



\end{abstract}